Project 1 Checkpoint

### 2015-02-28 project1/checkpoint/main.tex
Final tidying

Another entity to consider is the intersection itself. It's changing in time constantly, and those changes affect
every entity within the simulation. The most important parameters to consider here are the light timings. This
intersection's timings are time-based, rather than sensor-based, which can be seen in the data set provided with the
project description and is also available online \cite{ngsim}. We assume that the current light timing provided here
are optimal, but because we'll have a computer simulation in the end we have free reign in experimenting with these
values and observing the outcome. Perhaps the timings are \emph{not} optimal. Perhaps, our model grossly oversimplifies
some crucial interaction within the system. This is perhaps the most exciting part of the project.

@@

\subsubsection{Acceleration Model}
Bonneson \cite{bonneson} noted a property of car platoon acceleration: Acceleration decreases linearly with velocity.
The formulation he provides is a differential equation:
\[
\frac{dV}{dt} = A_{m} \left( 1 - \frac{V}{V_{m}} \right)
\]
Where $A_m$ and $V_m$ where the empirically calculated maximum acceleration and velocity of cars. The acceleration was
determined to be roughly 6.63 $\frac{ft}{s^2}$ and the maximum velocity simply the speed limit in the area (30 mph).
This equation is indeed solvable by hand and gives us a closed form for the position of car (assuming an initial velocity of 0):
\[
    p(t) = V_{m}t - \frac{V_{m}^2}{A_{m}} \left( 1 - e^{-\frac{A_{m}t}{V_{m}} } \right)
\]

Assuming all cars have length $L$, the $n$th car in the queue will have to travel $L(n-1)$ feet in order to cross the
intersection. Given our closed form above, in order to know how long it will be before a car passes we'll need to solve
\[
p(t) = L(n-1)
\]
Unfortunately, this must be done numerically. The function is fairly well behaved, and will in general have two roots
(only one positive). Using a positive initial guess (around 2 or so, judging by visual inspection), Newton's method
will converge within only a few steps. \ref{fig:acceleration} demonstrates some properties of this function
$0 = p(t) - L(n-1)$, one being
that as we move further back in the queue, the timing between cars approaches a constant. This is a result of cars
have more space to accerate to full speed.

@@

The car model will listen for this event as well. When a currently traveling car sees a red light, it will calculate
\emph{where} it needs to stop, and \emph{when} it will come to a stop. This will be done using a deceleration model
similar to \cite{deceleration}. If the light turns to green, the car instead needs to start accelerating (using one of
the models mentioned above). The model will adapt to light changes during movement by recalculating its estimated
exit time of the system.

\subsection{Simulation Input}
The two main sources of input into this simulation are the cars entering out section of Peachtree St, and the light
timing of the system. The combination of these two variables are what influence the smoothness
of traffic flow to the actual drivers in the system.

@@

The distribution of car arrival is very important, but has been difficult to model. The arrival intervals between 
northbound and southbound cars are provided in \ref{fig:northbound_arrival} and \ref{fig:southbound_arrival}.
Model fitting was attempted for both $\Gamma$ and $\Gamma$-mixture models on the raw data, but resulted in poor
fits. Furthermore, pruning outliers was attempted as well, but still yielded bad matches. The outlier critera used for
this determination is the common interquartile range approach, where values are rejected when
they lie outside of $median + 1.5IQR$, where $IQR$ is the interquartile range $Q_3 - Q_1$.

Given the difficulties with modeling we'll take a simpler approach and assume the ``true distribution`` is modeled by what was observed,
and simply plan to use the observed frequency distribution as our input distribution.

@@

\subsubsection{Data Analysis}
In the \texttt{data\_processing} folder, most of code used for distribution fitting and statistics can be found there.
The data analysis code was written in JavaScript, and runs completely in NodeJS \cite{nodejs}. NPM \cite{npm} was used
to manage dependencies and those dependencies can be seen in the \texttt{package.json} file. Of those packages, I have
authored the \texttt{chi-squared-test} module \cite{chiSquaredNpm}, and the \texttt{gamma-distribution}
\cite{gammaDistributionNpm}. Gulp \cite{gulp} was used a build tool for automating tasks (including building this paper).
To run some of the data analysis code I wrote, you'll need to install and run gulp. Instructions for doing so are
provided in the final section of this paper.

Alongside the NGSIM data set \cite{ngsim}, we are also dependent on the browser for much of our runtime configuration.
In particular, we rely on the random number generation capabilities of the browser for our entire source of randomness.
JavaScript provides a built-in implementation \cite{ecmascriptmath} \texttt{Math.random()}, which provides a uniformly
distributed random number between 0 and 1. The standard also specifies that the implementation is vendor-dependent.
This means that the "randomness" of \texttt{Math.random()} is dependent on the browser. Because of this, I've provided
a web page to calculate a simple $\chi^2$ statistic for a sample \cite{rngtester}. I've tested on Safari, Firefox, and
Chrome with generally solid results, despite the fact that their implementations are different
\cite{rngimplementation}. However, Hyland \cite{rngimplementation} does warn the reader about the caveats associated
with the native browser implementation of random numbers. However, these issues are not terribly important for a
simulation of this simplicity.

@@

Due to our inability to map a common distribution to the data gathered in the NGSIM study, we are using an empircal
distribution to model the input distribution of cars. Although this is an "easy" choice, it comes with the assumption
that both the measured data is representative of the true input distribution of cars, and that there is no measurement
error in the data collected.

@@

Since we want to gather metrics regarding the throughput of cars, we'll also provide a listener class
\texttt{ThroughputStatistics} that will listen for cars to complete their journey through the system. This will
be reported in the UI in real time.

@@

To build the application locally, you'll need to install a version of Node.js \cite{nodejs} at 0.10.0 or higher. This will be bundled
with npm. Since this project uses gulp for a task runner, you'll need to install it globally via
\texttt{npm install -g gulp}. Inside the \texttt{simulation} directory, you'll need to pull in packages with just
\texttt{npm install}. Once that installs, running \texttt{gulp} will run the build for you and generate a
\texttt{main.js} that's included in the \texttt{index.html} file. Opening this file in a browser should run the
simulation. To rerun the simulation, simply refresh the browser

>>>>

Another entity to consider is the intersection itself. It's changing in time constantly, and those changes affect
every entity within the simulation. The most important parameters to consider here are the light timings. This
intersection's timings are time-based, rather than sensor-based, which can be seen in the data set provided with the
project description and is also available online \cite{ngsim}. We assume that the current light timings provided here
are optimal, but because we'll have a computer simulation in the end we have free reign in experimenting with these
values and observing the outcome. Perhaps the timings are \emph{not} optimal. Perhaps, our model grossly oversimplifies
some crucial interaction within the system. This is perhaps the most exciting part of the project.

@@

\subsubsection{Acceleration Model}
Bonneson \cite{bonneson} noted a property of car platoon acceleration: Acceleration decreases linearly with velocity.
The formulation he provides is a differential equation:
\[
\frac{dV}{dt} = A_{m} \left( 1 - \frac{V}{V_{m}} \right)
\]
where $A_m$ and $V_m$ are the empirically calculated maximum acceleration and velocity of cars. The acceleration was
determined to be roughly 6.63 $\frac{ft}{s^2}$ and the maximum velocity simply the speed limit in the area (30 mph).
This equation is indeed solvable by hand and gives us a closed form for the position of car (assuming an initial velocity of 0):
\[
    p(t) = V_{m}t - \frac{V_{m}^2}{A_{m}} \left( 1 - e^{-\frac{A_{m}t}{V_{m}} } \right)
\]

Assuming all cars have length $L$, the $n$th car in the queue will have to travel $L(n-1)$ feet in order to cross the
intersection. Given our closed form above, in order to know how long it will be before a car passes we'll need to solve
\[
p(t) = L(n-1)
\]
Unfortunately, this has no closed form and must be done numerically. The function is fairly well behaved, and will in general have two roots
(only one positive). Using a positive initial guess (around 2 or so, judging by visual inspection), Newton's method
will converge within only a few steps. Figure \ref{fig:acceleration} demonstrates some properties of this function
$0 = p(t) - L(n-1)$, one being
that as we move further back in the queue, the timing between cars approaches a constant. This is a result of cars
have more space to accelerate to full speed.

@@

Furthermore, the queue at the intersection will need to listen for this event as well. In essense, when the light
changes, the intersection will need to switch between holding cars and actually passing them through. If the light has
turned to green, it will need to one-by-one discharge cars based on the above acceleration model. If the light is red,
the queue will need to appropriate insert the car to await the car's turn to pass through the intersection.

\subsection{Simulation Input}
The two main sources of input into this simulation are the cars entering our section of Peachtree St, and the light
timing of the system. The combination of these two variables are what influence the smoothness
of traffic flow to the actual drivers in the system.

@@

The distribution of car arrival is very important, but has been difficult to model. The arrival intervals between 
northbound and southbound cars are provided in \ref{fig:northbound_arrival} and \ref{fig:southbound_arrival}.
Model fitting was attempted for both $\Gamma$ and $\Gamma$-mixture models on the raw data, but resulted in poor
fits (verified by $\chi^2$ goodness-of-fit tests. Furthermore, pruning outliers was attempted as well, but still
yielded bad matches. The outlier critera used for
this determination is the common interquartile range approach, where values are rejected when
they lie outside of $median + 1.5IQR$, where $IQR$ is the interquartile range $Q_3 - Q_1$.

Given the difficulties with modeling we'll take a simpler approach and assume the true distribution is modeled by what was observed,
and simply plan to use the observed frequency distribution as our input distribution.

@@

\subsubsection{Data Analysis}
In the \texttt{data\_processing} folder, most of code used for distribution fitting and statistics can be found there.
The data analysis code was written in JavaScript, and runs completely in NodeJS \cite{nodejs}. NPM \cite{npm} was used
to manage dependencies and those dependencies can be seen in the \texttt{package.json} file. Of those packages, I have
authored the \texttt{chi-squared-test} module \cite{chiSquaredNpm}, and the \texttt{gamma-distribution}
\cite{gammaDistributionNpm}. Gulp \cite{gulp} was used as a build tool for automating tasks (including building this paper).
To run some of the data analysis code I wrote, you'll need to install and run gulp. Instructions for doing so are
provided in the final section of this paper.

Alongside the NGSIM data set \cite{ngsim}, we are also dependent on the browser for much of our runtime configuration.
In particular, we rely on the random number generation capabilities of the browser for our entire source of randomness.
JavaScript provides a built-in implementation \cite{ecmascriptmath} \texttt{Math.random()}, which provides a uniformly
distributed random number between 0 and 1. The standard specifies that the implementation is vendor-dependent.
This means that the ``randomness'' of \texttt{Math.random()} is dependent on the browser. Because of this, I've provided
a web page to calculate a simple $\chi^2$ statistic for a sample of \texttt{Math.random()} \cite{rngtester}. I've tested
\texttt{Math.random()} on Safari, Firefox, and Chrome with generally solid results, despite the fact that their implementations
are different \cite{rngimplementation}. However, Hyland \cite{rngimplementation} does warn the reader about the caveats associated
with the native browser implementation of random numbers. However, these issues are not terribly important for a
simulation of this simplicity.

@@

Due to our inability to map a common distribution to the data gathered in the NGSIM study, we are using an empirical
distribution to model the input distribution of cars. Although this is an ``easy'' choice, it comes with the assumption
that both the measured data is representative of the true input distribution of cars, and that there is no measurement
error in the data collected.

@@

Since we want to gather metrics regarding the throughput of cars, we'll also provide a listener class
\texttt{StatCollection} that will listen for cars to complete their journey through the system. This will
be reported in the UI in real time.

@@

To build the application locally, you'll need to install a version of Node.js \cite{nodejs} at 0.10.0 or higher. This will be bundled
with npm. Since this project uses gulp for a task runner, you'll need to install it globally via
\texttt{npm install -g gulp}. Inside the \texttt{simulation} directory, you'll need to pull in packages with just
\texttt{npm install}. Once that installs, running \texttt{gulp} will run the build for you and generate a
\texttt{main.js} that's included in the \texttt{index.html} file. Opening this file in a browser should run the
simulation. To rerun the simulation, simply refresh the page.
